---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r echo=TRUE}
setwd("C:/Users/210067772/Coursera/data")
data<- read.csv("activity.csv", colClasses=c("numeric", "character", "numeric"))
```

## What is mean total number of steps taken per day?

Let's ignore the missing values in the dataset.

1. Histogram of the total number of steps taken each day

```{r echo=TRUE}
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)

hist(total.steps, 
     main="Histogram of the Total Number of Steps taken per day (NA removed)", 
     xlab="Total number of steps per day", 
     ylab="Frequency",
     border="blue", 
     col="green",
     xlim=c(0,25000),
     ylim=c(0, 20),
     las=1, 
     breaks=seq(from=0, to=25000, by=2500)
     )
```

2. Calculate and report the mean and median total number of steps taken per day


```{r echo=TRUE}
mean(total.steps)
median(total.steps)
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r echo=TRUE}
av.steps <- aggregate(x = data$steps, by = list(data$interval), FUN = "mean", na.rm=TRUE)
names(av.steps) <- c("Interval", "Average")
plot(av.steps$Interval,
     av.steps$Average, 
     type = "l", 
     col = "blue",
     xlab = "Interval (min)", 
     ylab = "Average Number of Steps",
     main="Time-series of the average number of steps per intervals\n(NA removed)")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
av.steps[which.max(av.steps$Average),] 
```
## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r echo=TRUE}
count.NA <- sum(is.na(data$steps)) #2304
```
2. Devise a strategy for filling in all of the missing values in the dataset. I decided to use the mean of the sample.


```{r echo=TRUE}
# Find the NA positions
pos.NA <- which(is.na(data$steps))

# create a vector with the mean of the steps, we'll fill the missing values with this vector
fill.NA <- rep(mean(data$steps, na.rm=TRUE), times=length(pos.NA))
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r echo=TRUE}
data[pos.NA, "steps"] <- fill.NA
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r echo=TRUE}
new.total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
hist(new.total.steps, 
     main="Histogram of the Total Number of Steps taken per day\n (NA replaced by the mean of the sample)", 
     xlab="Total number of steps per day", 
     ylab="Frequency",
     border="blue", 
     col="green",
     xlim=c(0,25000),
     ylim=c(0, 25),
     las=1, 
     breaks=seq(from=0, to=25000, by=2500)
)

mean(new.total.steps) #10766.19
median(new.total.steps) #10766.19
```





## Are there differences in activity patterns between weekdays and weekends?




